---
title: 'Week 3: R Markdown Presentation & Plotly'
author: "Ingrid Tobar"
date: "7/9/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Instructions

- Create a web page presentation using R Markdown that features a plot created with Plotly.
- Host the webpage on either GitHub Pages, RPubs, or NeoCities. 
- The webpage must contain the date when it was created and a plot created with Plotly. 
- The next slide shows a **Grouped Box Plot** of the diamond dataset, showing price (y axis) vs clarity (x axis), grouped by cut (Fair, Good, Very Good, Premium, Ideal)

## Grouped Box Plot

```{r message=FALSE, warning=FALSE}
library(plotly)
p <- plot_ly(ggplot2::diamonds, x = ~cut, y = ~price, color = ~clarity, type = "box") %>%
  layout(boxmode = "group")
p
```

<font size="3">**Date created:** July 9, 2019</font>

<font size="3">**Author:** Ingrid Tobar</font>

## End of Slide Show


This is the last slide of the slide show

Thanks!